Nested Sampling with Multiple Objective Functions

### 2018-06-02 doc/ms.tex
Partitoin functions keyword
Working on the abstract

\DeclareUnicodeCharacter{00A0}{ }
\newcommand{\xx}{\boldsymbol{\theta}}
\newcommand{\dx}{d\boldsymbol{\theta}}
\newcommand{\data}{\boldsymbol{D}}
\newcommand{\II}{\boldsymbol{I}}

@@

% Abstract (Do not use inserted blank lines, i.e. \\)
\abstract{We present a generalization of the Nested Sampling algorithm
that estimates
the partition function $Z(\beta_1, \beta_2)$ for a family of
probability distributions indexed by two hyperparameters, $\beta_1$ and
$\beta_2$. Problems like this arise frequently in statistical mechanics and
occasionally in Bayesian inference.
We demonstrate the algorithm on three example problems.
The first is a toy problem whose structure allows us
to accurately calculate $Z(\beta_1, \beta_2)$ numerically using non-Monte-Carlo
methods. With this example, we empirically investigate the accuracy of TwinPeaks
improves as a function of the number of particles used, $N$. We also
calculate the partition function for a collection of atoms with a potential
that is intermediate between a Lennard-Jones form and a `polymer' form.
Finally, we present a Bayesian image reconstruction example,
where we can infer the value of a hyperparameter even though its prior
distribution contains an `intractable' normalizing constant.
An implementation in C++ is available at
\url{https://github.com/eggplantbren/TwinPeaks2018}.}

% Keywords: add 3 to 10 keywords
\keyword{nested sampling; bayesian computation; statistical mechanics;
         sequential monte carlo; multi-objective optimization}

@@

Nested Sampling \citep[NS][]{skilling} is popular general
Monte Carlo algorithm that can solve a wide range of problems in Bayesian
computation and statistical mechanics
\citep{2009arXiv0906.3544P, 2014PhRvE..89b2302P, 2015arXiv150303404B}.
Its key strength its ability to cope with phase transitions which cause
problems for many other methods \citep{skilling}.
In a Bayesian inference problem with unknown parameters denoted collectively
by a vector $\xx$, the
posterior distribution for the parameters given data $\data$ (and prior
information $\II$) is:
\begin{eqnarray}
p(\xx | \data, \II) &=&
\frac{p(\xx | \II)p(\data | \xx, \II)}{p(\data | \II)}\\
&=& \frac{\pi(\xx)L(\xx)}{Z}
\end{eqnarray}
where $\pi(\xx)$ is the prior distribution, $L(\xx)$ is the likelihood
function, and $Z$ is the normalising constant, usually known as the
`marginal likelihood' or the `evidence':
\begin{eqnarray}
Z &=& \int \pi(\xx) L(\xx) \, \dx.\label{eqn:evidence}
\end{eqnarray}

>>>>

\DeclareUnicodeCharacter{00A0}{ }
\newcommand{\xx}{\boldsymbol{\theta}}
\newcommand{\dx}{d\boldsymbol{\theta}}
\newcommand{\data}{D}
\newcommand{\II}{I}

@@

% Abstract (Do not use inserted blank lines, i.e. \\)
\abstract{We present a generalization of Skilling's Nested Sampling algorithm
that estimates
the partition function $Z(\beta_1, \beta_2)$ for a family of
probability distributions indexed by two hyperparameters, $\beta_1$ and
$\beta_2$. Problems like this arise frequently in statistical mechanics and
occasionally in Bayesian inference.
The algorithm is based on an interpretation of Nested Sampling in terms of
Sequential Monte Carlo, which produces unbiased estimates of $Z$ and therefore
allows parallel runs to be trivially combined.
We demonstrate the algorithm on three example problems:
(i)
a toy problem whose structure allows us
to accurately calculate $Z(\beta_1, \beta_2)$ numerically using non-Monte-Carlo
methods, so we can compute the actual error in the reconstructed function
$Z(\beta_1, \beta_2)$; (ii) a collection of atoms with a potential
that is intermediate between a Lennard-Jones form and a `polymer' form; and
(iii) A Bayesian image reconstruction example where the prior for the image
is defined conditionally on a hyperparameter, but where this prior contains
an intractable normalizing constant.
A free software implementation in C++ and Python is available at
\url{https://github.com/eggplantbren/TwinPeaks2018}.}

% Keywords: add 3 to 10 keywords
\keyword{nested sampling; bayesian computation; statistical mechanics;
         partition functions;
         sequential monte carlo; multi-objective optimization}

@@

Nested Sampling \citep[NS][]{skilling} is popular general
Monte Carlo algorithm that can solve a wide range of problems in Bayesian
computation and statistical mechanics
\citep{2009arXiv0906.3544P, 2014PhRvE..89b2302P, 2015arXiv150303404B}.
Its key strength its ability to cope with phase transitions which cause
problems for many other methods \citep{skilling}.
In a Bayesian inference problem with unknown parameters denoted collectively
by a vector $\xx$, the
posterior distribution for the parameters given a data proposition $\data$ and
a `prior information' proposition $\II$ is
information $\II$) is:
\begin{eqnarray}
p(\xx | \data, \II) &=&
\frac{p(\xx | \II)p(\data | \xx, \II)}{p(\data | \II)}\\
&=& \frac{\pi(\xx)L(\xx)}{Z}
\end{eqnarray}
where $\pi(\xx)$ is the prior distribution, $L(\xx)$ is the likelihood
function, and $Z$ is the normalising constant, usually known as the
`marginal likelihood' or the `evidence':
\begin{eqnarray}
Z &=& \int \pi(\xx) L(\xx) \, \dx.\label{eqn:evidence}
\end{eqnarray}
